Illuminating a Dark Lens : A Type Ia Supernova Magnified by \\ the \changeR{Frontier Fields} Galaxy Cluster Abell 2744

### 2015-05-22 snTomas.tex
final tweaks during arXiv submission

\begin{abstract}
\changeG{
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744 ($z=0.308$).
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of $\mu_{\rm
obs}=2.03\pm0.29$.  This observed magnification provides a rare
opportunity for a direct empirical test of galaxy cluster lens models.
Here we test 17 lens models, 13 of which were generated before the SN
magnification was known, qualifying as pure ``blind tests''.  The
models are collectively fairly accurate: 8 of the models deliver
median magnifications that are consistent with the measured $\mu$ to
within 1$\sigma$.  However, there is a subtle systematic bias: the
significant disagreements all involve models {\it overpredicting} the
magnification. We evaluate possible causes for this mild bias, and
find no single physical or methodological explanation to account for
it.  We do find that model accuracy can be improved to some extent
with stringent quality cuts on multiply-imaged systems, such as
requiring that a large fraction have spectroscopic redshifts.  In
addition to testing model accuracies as we have done here, Type Ia SN
magnifications could also be used as inputs for future lens models of
Abell 2744 and other clusters, providing valuable constraints in
regions where traditional strong- and weak-lensing information is
unavailable.}
\end{abstract}

@@

We thank the Hubble Frontier Fields team at STScI for their
substantial efforts to make the HFF program successful.  In
particular, thanks are due to Matt Mountain for the allocation of
discretionary orbits for the HFF program; to Jennifer Lotz, Norman
Grogin and Patricia Royle for accommodations in strategy and
implementation to make the FrontierSN program possible; to Anton
Koekemoer for HFF data processing support; and to Dan Coe for curating the
excellent and accessible lens model comparison tools.  We also thank
the CLASH team, led by Marc Postman, for observations, catalogs, and
high level science products that were of significant value for this
analysis.  Thanks to Jonatan Selsing for helpful comments on the
manuscript.

@@

{\it Facilities:} \facility{HST (WFC3)}
\smallskip

>>>>

\begin{abstract}
{
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744 ($z=0.308$).
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of $\mu_{\rm
obs}=2.03\pm0.29$.  This observed magnification provides a rare
opportunity for a direct empirical test of galaxy cluster lens models.
Here we test 17 lens models, 13 of which were generated before the SN
magnification was known, qualifying as pure ``blind tests''.  The
models are collectively fairly accurate: 8 of the models deliver
median magnifications that are consistent with the measured $\mu$ to
within 1$\sigma$.  However, there is a subtle systematic bias: the
significant disagreements all involve models {\it overpredicting} the
magnification. We evaluate possible causes for this mild bias, and
find no single physical or methodological explanation to account for
it.  We do find that model accuracy can be improved to some extent
with stringent quality cuts on multiply-imaged systems, such as
requiring that a large fraction have spectroscopic redshifts.  In
addition to testing model accuracies as we have done here, Type Ia SN
magnifications could also be used as inputs for future lens models of
Abell 2744 and other clusters, providing valuable constraints in
regions where traditional strong- and weak-lensing information is
unavailable.}
\end{abstract}

@@

We thank the Hubble Frontier Fields team at STScI for their
substantial efforts to make the HFF program successful.  In
particular, thanks are due to Matt Mountain for the allocation of
discretionary orbits for the HFF program; to Jennifer Lotz, Norman
Grogin and Patricia Royle for accommodations in strategy and
implementation to make the FrontierSN program possible; to Anton
Koekemoer for HFF data processing support.  We also thank
the CLASH team, led by Marc Postman, for observations, catalogs, and
high level science products that were of significant value for this
analysis.  Thanks to Jonatan Selsing for helpful comments on the
manuscript.

@@

{\it Facilities:} \facility{HST (WFC3)}
\pagebreak
